Coherent optical association of a single molecule

### 2020-10-16 paper.tex
Update on Overleaf.

Diverse species of fully quantum controlled  ultracold molecules are desired for a  wide variety of applications including precision measurements~\cite{Nick_and_Ivan2017}, quantum simulations~\cite{Yao2018}, quantum  information processing~\cite{DeMille2002, Ni2018}, and studies of ultracold chemistry~\cite{Bohn2017,Bala2016,Hu1111}.
While many innovative approaches demonstrated in the last few years have directly cooled different species of diatomic or polyatomic molecules below 1~mK~\cite{Norrgard2016, Mitra1366}, the coldest and the highest phase-space-density gas to date in an ensemble~\cite{Demarco2018} or as individuals~\cite{Zhang2020}  have been achieved through the association of ultracold atoms.
% (any place to cite  molecular ion work?)
% cite recent doyle group result on cooling symmetric top molecule? FOR SURE, and other laser cooling works and electro-optical-cooling! Citations are not complete. Add them as you see fit.

@@

As previously discussed, the ratio of the Raman Rabi rate to the scattering rate is proportional to the $ \Omega_a/\Omega_m $ ratio. The smaller this ratio is, the farther the laser needs to be detuned to achieve the same Raman Rabi rate to scattering rate ratio, which lowers the Raman Rabi rate. In addition to reducing how far the laser needs to be detuned from the transition, a larger $ \Omega_a/\Omega_m$ ratio also relaxes the intensity stability requirement, a key potential technical limitation. The position of the two photon resonance depends on the laser power predominantly through the AC Stark shift on the molecular state, $ \Omega_m^2 / 2\Delta $ \footnote{There is an additional factor of 2 to account for the Stark shift of both beams.}. The ratio of the AC Stark shift to the Raman Rabi frequency is $ \Omega_m / \Omega_a $, Thus, the laser intensity needs to be stabilized to better than the inverse of this ratio to fluctuate by less than a linewidth in a coherent process. The ratio, $ \Omega_a/\Omega_m $ can be changed through the choice of initial and final states. In particular, it can be made larger by increasing the coupling of the atomic state with the excited molecular state. Due to the small size of the molecular wavefunction, the coupling between the ground atomic state and the excited molecular state is approximately proportional to the value of the relative atomic wavefunction at short distance within the molecular potential. In addition to the confinement, this value is related to the interaction between the two atoms. For states with a large scattering length (positive or negative), the phase shift in the relative wavefunction between the atoms can significantly increase the short range wavefunction (Fig.~\ref{f-theory}C). The increase in the coupling is proportional to (quote/cite Olive's equation?). For our system, we denote the possible hyperfine states of the atoms as $ \ket{\uparrow_{\text{Cs}}} = \ket{F=4,m_F=4}_{\text{Cs}}, \ket{\downarrow_{\text{Cs}}} = \ket{F=3,m_F=3}_{\text{Cs}}, \ket{\uparrow_{\text{Na}}} = \ket{F=2,m_F=2}_{\text{Na}}, $ and $ \ket{\downarrow_{\text{Na}}} = \ket{F=1,m_F=1}_{\text{Na}}$. Among the stable spin combinations, $\ket{\uparrow_{\text{Na}}\uparrow_{\text{Cs}}}$ and $\ket{\downarrow_{\text{Na}}\downarrow_{\text{Cs}}} $ both have small scattering lengths of $ 30.4a_0 $, and $ 13.7a_0 $ respectively, but the $\ket{\uparrow_{\text{Na}} \downarrow_{\text{Cs}}} $ combination has a large and negative scattering length of $ a(\uparrow_{\text{Na}} \downarrow_{\text{Cs}}) = -693.8a_0 $ (interaction shift $\approx$ binding?)~\cite{Hood2019}. In addition to the increased atomic coupling, $ \Omega_a $, with the $\ket{\uparrow_{\text{Na}} \downarrow_{\text{Cs}}}$ hyperfine combination, coupled channel calculations show that the most weakly bound molecular state that is predominantly in this spin combination also has reduced coupling, $ \Omega_m $, with the excited state when compared with the $\ket{\uparrow_{\text{Na}}\uparrow_{\text{Cs}}}$ and $\ket{\downarrow_{\text{Na}}\downarrow_{\text{Cs}}} $ bound states. Therefore, using an initial $\ket{\uparrow_{\text{Na}} \downarrow_{\text{Cs}}}$ hyperfine combination results in a $ \Omega_a/\Omega_m$ ratio of about 0.05 instead of a ratio of about 0.003 with the other combinations.

@@

In conclusion, we have formed a weakly bound NaCs molecule in an optical tweezer via an optical Raman transfer. A theoretical investigation including all excited states of $ c^3\Sigma^+(\Omega = 1)$ and coupled channel ground state wavefunctions indicated better transfer efficiency using a deeply bound intermediate state and the $ \ket{\uparrow_{\text{Na}} \downarrow_{\text{Cs}} } $ hyperfine state combination. Using these theoretical insights, we located the weakly bound state and coherently associated the atoms into the weakly bound molecule. Our transfer efficiency is limited by an unknown scattering source resulting in measured scattering rates over 10 times larger than theory predictions. Despite this limitation, the transfer efficiency may be further improved by increasing the $ \Omega_a/\Omega_m$ ratio by exploring the possibility of driving to more deeply bound states. There may also be a better choice of single photon detuning to increase the Raman Rabi frequency, since our location results in about 30\% cancellation of the Raman Rabi frequency due to an offset of unknown origin. (Do we want to mention also technical improvements as another way to potentially improve our signal? Though I guess that was mentioned directly in the paragraph above...) Nevertheless, our technique, with the right choice of parameters, can be used to form a more diverse set of molecular species, since it does not rely on a magnetic Feshbach resonance or a narrow excited state (More sentences here to discuss scientific outlook, but without repeating the introduction...? \textcolor{blue}{Yes, on topics related to single molecules for exciting science!!})

>>>>

Diverse species of fully quantum controlled  ultracold molecules are desired for a  wide variety of applications including precision measurements~\cite{Nick_and_Ivan2017, PhysRevA.101.042504} \textcolor{red}{KW: added a citation here for some exotic atomic/molecular species for precision measurements. Can delete this text, if you think it's appropriate, or otherwise remove this reference.}, quantum simulations~\cite{Yao2018}, quantum  information processing~\cite{DeMille2002, Ni2018}, and studies of ultracold chemistry~\cite{Bohn2017,Bala2016,Hu1111}.
While many innovative approaches demonstrated in the last few years have directly cooled different species of diatomic or polyatomic molecules below 1~mK~\cite{Norrgard2016, Mitra1366}, the coldest and the highest phase-space-density gas to date in an ensemble~\cite{Demarco2018} or as individuals~\cite{Zhang2020}  have been achieved through the association of ultracold atoms.
% (any place to cite  molecular ion work?)
% cite recent doyle group result on cooling symmetric top molecule? FOR SURE, and other laser cooling works and electro-optical-cooling! Citations are not complete. Add them as you see fit.

@@

As previously discussed, the ratio of the Raman Rabi rate to the scattering rate is proportional to the $ \Omega_a/\Omega_m $ ratio. The smaller this ratio is, the farther the laser needs to be detuned to achieve the same Raman Rabi rate to scattering rate ratio, which lowers the Raman Rabi rate. In addition to reducing how far the laser needs to be detuned from the transition, a larger $ \Omega_a/\Omega_m$ ratio also relaxes the intensity stability requirement, a key potential technical limitation. The position of the two photon resonance depends on the laser power predominantly through the AC Stark shift on the molecular state, $ \Omega_m^2 / 2\Delta $ \footnote{There is an additional factor of 2 to account for the Stark shift caused by both beams.}. The ratio of the AC Stark shift to the Raman Rabi frequency is $ \Omega_m / \Omega_a $, Thus, the laser intensity needs to be stabilized to better than the inverse of this ratio to fluctuate by less than a linewidth in a coherent process. The ratio, $ \Omega_a/\Omega_m $ can be changed through the choice of initial and final states. In particular, it can be made larger by increasing the coupling of the atomic state with the excited molecular state. Due to the small size of the molecular wavefunction, the coupling between the ground atomic state and the excited molecular state is approximately proportional to the value of the relative atomic wavefunction at short distance within the molecular potential. In addition to the confinement, this value is related to the interaction between the two atoms. For states with a large scattering length (positive or negative), the phase shift in the relative wavefunction between the atoms can significantly increase the short range wavefunction (Fig.~\ref{f-theory}C). The increase in the coupling is proportional to (quote/cite Olive's equation?). For our system, we denote the possible hyperfine states of the atoms as $ \ket{\uparrow_{\text{Cs}}} = \ket{F=4,m_F=4}_{\text{Cs}}, \ket{\downarrow_{\text{Cs}}} = \ket{F=3,m_F=3}_{\text{Cs}}, \ket{\uparrow_{\text{Na}}} = \ket{F=2,m_F=2}_{\text{Na}}, $ and $ \ket{\downarrow_{\text{Na}}} = \ket{F=1,m_F=1}_{\text{Na}}$. Among the stable spin combinations, $\ket{\uparrow_{\text{Na}}\uparrow_{\text{Cs}}}$ and $\ket{\downarrow_{\text{Na}}\downarrow_{\text{Cs}}} $ both have small scattering lengths of $ 30.4a_0 $, and $ 13.7a_0 $ respectively, but the $\ket{\uparrow_{\text{Na}} \downarrow_{\text{Cs}}} $ combination has a large and negative scattering length of $ a(\uparrow_{\text{Na}} \downarrow_{\text{Cs}}) = -693.8a_0 $ (interaction shift $\approx$ binding?)~\cite{Hood2019}. In addition to the increased atomic coupling, $ \Omega_a $, with the $\ket{\uparrow_{\text{Na}} \downarrow_{\text{Cs}}}$ hyperfine combination, coupled channel calculations show that the most weakly bound molecular state that is predominantly in this spin combination also has reduced coupling, $ \Omega_m $, with the excited state when compared with the $\ket{\uparrow_{\text{Na}}\uparrow_{\text{Cs}}}$ and $\ket{\downarrow_{\text{Na}}\downarrow_{\text{Cs}}} $ bound states. Therefore, using an initial $\ket{\uparrow_{\text{Na}} \downarrow_{\text{Cs}}}$ hyperfine combination results in a $ \Omega_a/\Omega_m$ ratio of about 0.05 instead of a ratio of about 0.003 with the other combinations.

@@

In conclusion, we have formed a weakly bound NaCs molecule in an optical tweezer via an optical Raman transfer. A theoretical investigation including all excited states of $ c^3\Sigma^+(\Omega = 1)$ and coupled channel ground state wavefunctions indicated better transfer efficiency using a deeply bound intermediate state and the $ \ket{\uparrow_{\text{Na}} \downarrow_{\text{Cs}} } $ hyperfine state combination. Using these theoretical insights, we located the weakly bound state and coherently associated the atoms into the weakly bound molecule. Our transfer efficiency is limited by an unknown scattering source resulting in measured scattering rates over 10 times larger than theory predictions. Despite this limitation, the transfer efficiency may be further improved by increasing the $ \Omega_a/\Omega_m$ ratio by exploring the possibility of driving to more deeply bound states. There may also be a better choice of single photon detuning to increase the Raman Rabi frequency, since our location results in about 30\% cancellation of the Raman Rabi frequency due to an offset of unknown origin. (Do we want to mention also technical improvements as another way to potentially improve our signal? Though I guess that was mentioned directly in the paragraph above...) Nevertheless, our technique, with the right choice of parameters, can be used to form a more diverse set of molecular species, since it does not rely on a magnetic Feshbach resonance or a narrow excited state. The formation of a weakly bound molecule is also a key step in forming rovibrational ground state molecules. Combined with real time rearrangement \cite{Barredo2016, Endres2016}, defect free arrays of highly controlled molecules can be formed in arbitrary geometries. These arrays are a promising and flexible platform for quantum simulation and quantum computing applications.
